---
title: "Class 1"
author: "Christian"
date: "2025-10-16"
output: html_document
---

## R for data science class 1
 Dr. Pacifique
 2025-09-14
 Important definitions in data science
 Statistics
 Modelling
 Computation
 Calculus
## Today’s class,we will study
 Installation of R
 Installation of R studio
 Calculator in r
 Data importation in R
 Data description using numerical measures and graphs.
## Data structure (Vectors, Factors, Lists, Data frame, matrix and arrays)
## Vector: The foundational R data structure
## Numeric

```{r}
x <- c(0.5, 0.6)
 age<-c(20,35,32,29)
 summary(age)
```
```{r}
 age<-c(35,24,18,24)
 mean(age)
```
### Logical
```{r}
x <- c(TRUE, FALSE)
 x <- c(T, F)
```

### Character

```{r}
x<-c("a","b","c")
 class<-c("M","F","F","M")
```

### Integer
```{r}
x<-9:29
```

### Complex
```{r}
 x<-c(1+0i,2+4i)
```

### Matrixandoperationwithinmatrices
```{r}
x<-matrix(1:6,nrow=2,ncol=3)
 x
```
```{r}
(y<-matrix(1:6,nrow=3,ncol=2))
```
```{r}
 x%*%y
```
### List

```{r}
x<-list(1,"a",TRUE,1+4i)
 x
```

### Factor

```{r}
 x<-factor(c("yes","yes","no","yes","no"))
```

### CreateavectorwithNAsinit
```{r}
x<-c(1,2,NA,10,3)
```

ReturnalogicalvectorindicatingwhichelementsareNA

```{r}
 is.na(x)
```

```{r}
 bad<-is.na(x)
 print(bad)
```

```{r}
x[!bad]
```

WhatiftherearemultipleRobjectsandyouwanttotakethesubsetwithnomissing
 valuesinanyofthoseobjects?
 
```{r}
 x<-c(1,2,NA,4,NA,5)
 y<-c("a","b",NA,"d",NA,"f")
 good<-complete.cases(x,y)
 good
```
 
###  Coercion
 Ifcharacterispresent,inavector,Rconverteverythinginthevectortocharacter
 strings.
 Ifavectoronlycontainslogicalandnumbers,Rwillconvertthelogicaltonumbers,
 Everytruebecomesa1,andeveryFALSEbecomes0
```{r}
sum(c(TRUE,TRUE,FALSE,FALSE,FALSE))
```

