```{r}
#6.15 Obtain the stratified importance sampling estimate in Example 6.14
# and compare it with the result of Example 6.11.
```
1.Crude Monte Carlo (Example 6.11)

```{r}
set.seed(1)

n <- 100000
x <- rnorm(n)
I_MC <- mean(x^2 * (x >= 1))

I_MC

```
2.Importance Sampling (Example 6.14)

```{r}
set.seed(1)

n <- 100000
u <- runif(n)
x <- qnorm(u * (1 - pnorm(1)) + pnorm(1))
I_IS <- (1 - pnorm(1)) * mean(x^2)

I_IS

```
3.Stratified Importance Sampling (Example 6.15)

```{r}
# Divide samples equally among 3 strata.

set.seed(1)
n <- 100000
n1 <- n2 <- n3 <- n/3

# step 1: Compute stratum probabilities.

p1 <- pnorm(2) - pnorm(1)
p2 <- pnorm(3) - pnorm(2)
p3 <- 1 - pnorm(3)

# Step 2: Function to sample truncated normal

rtrunc <- function(n, a, b) {
  u <- runif(n)
  qnorm(u * (pnorm(b) - pnorm(a)) + pnorm(a))
}
# Step 3: Sample each stratum

# Forces samples to come from each part of the tail.

x1 <- rtrunc(n1, 1, 2)
x2 <- rtrunc(n2, 2, 3)
x3 <- rtrunc(n3, 3, Inf)
# Step 4: Combine estimates

I_SIS <- p1 * mean(x1^2) +
         p2 * mean(x2^2) +
         p3 * mean(x3^2)

I_SIS







```
In conclusion:

1.The crude Monte Carlo estimator has high variance because most normal samples fall below 1 and contribute zero.
2.Importance sampling improves efficiency by sampling from a truncated normal distribution on 
[1,∞).
3.Stratified importance sampling further reduces variance by dividing the tail region into strata and sampling each separately, ensuring better coverage of the integration domain.


